Accurate binding of calcium to phospholipid bilayers by effective inclusion of electronic polarization

### 2017-07-28 Manuscript/proto-manuscript.tex
Introduction updated

% Here I write pseudo-article statements that will make the main argument.
% Beautiful polished sentences will be formed only afte we agree on these basic things
\section{Introduction}
Cation interactions with cellular membranes play a key role in several biological processes,
like in signal propagation in neurons and vesicle fusion.
Cation interactions with lipid bilayers serving as a model for cellular
membranes are thus widely studied with experimental \cite{cevc90,tocanne90,binder02,pabst07,uhrikova08}
and theoretical methods \cite{Berkowitz12,??}. General conclusion from experimental
studies has been that multivalent ions and lithium have weak specific binding in
phopholipid bilayers, while other monovalent ions do not essentially
bind \cite{cevc90,tocanne90,seelig90,binder02}. The presence of anoinic lipids, like PS or PG,
increase the concentration close to the bilayer and thus the amount of bound
ions, but do not affect the specific binding constant \cite{seelig90}.
The binding details, like binding sites and stoichiometry are not yet fully
resolved but interpretation of NMR and scattering experiments suggest that one
Ca2+ interacts mainly with the choline groups \cite{hauser76,hauser78,herbette84} of two
phospholipid molecules \cite{altenbach84}.

Classical molecular dynamics simulations have potential to solve the ion binding details
in lipid bilayers and reveal its relevance to various biological
problems \cite{bockmann03,bockmann04,??}. However, the available
molecular dynamics simulation models have a strong tendency to overestimate
cation binding on zwitterionic bilayers and none of them can reproduce the
experimental data with the accuracy required for the interpretation of
experiments \cite{catte16}. The overestimated specific cation binding also
makes zwitterionic bilayers effectively positively charged, which could
potentially lead to significant artefacts in applications.
Thus, there is a vast demand for classical molecular dynamics simulation
model which correctly reproduces cation binding in lipid bilayers.

The lack of electronic polarizability from the classical MD simulation
models is a potential source of artefact, which could lead to overbinding
of cations. 
%Almost all lipid bilayer simulations reported in the literature 
%are done with force fields that represent interactions between atoms
%with pairwise additive empirical potential and exclude the electronic 
%polarizability \cite{SOME_review, FF_papers}.
The issue has been considered higly relevant since the early days of
lipid bilayer simulations and pioneering simulation studies scaled
the partial charges to effectively include electronic
polarizability \cite{jonsson86,egberts94}. The explicit inclusion of
electronic polarizability has been, however, turned to be 
practically complicated and implementations for lipids are rare \cite{chowdhary13}.

In this work we show that the cation overbinding in classical MD simulations
can be corrected by including electronic polarizability by using effective continuum
correction (ECC)~\cite{leontyev11} for polar region of zwitterionic lipid molecules.
This is essentially physicalle well justified version of partial charge scaling
implemented in early days of lipid and surfctant simulations \cite{jonsson86,egberts94}. 
The approach has been previously shown in to improve bulk performace of
ion models against neutron scattering data \cite{Jungwirth2015,kohagen14,kohagen16}.
The better bulk behaviour was not, however, sufficient to correct binding in lipid
bilayers \cite{catte16}. 
%motivation for its application to zwitterionic lipids like POPC.

@@

\subsection{Electronic continuum correction for lipid bilayers}
According to the electronic continuum correction (ECC)\cite{leontyev11}, electronic
polarizability can be included in classical MD simulations by
placing all particles into a homogeneous dielectric continuum 
with a dielectric constant $\epsilon _{el}$, 
which is the electronic part of the dielectric constant of 
the media \cite{leontyev11}. Measurements of high frequency 
dielectric constant gives values of approximately $\epsilon _{el} \approx 2$ 
for almost any biomaterial \cite{some_original_work, leontyev11}. % that exists in living cells. 
Such a continuum can be easily included in standard MD simulation by
a formal transformation of partial charges 
\begin{equation}
  Q^{ECC} = f_q \cdot Q
\end{equation}
with a constant scaling factor $f_q = \epsilon _{el} ^{-1/2}$ 
effectively representing the newly introduced electronic continuum. 
The value measured for water, $\epsilon _{el} = 1.78$, gives 
a scaling factor of $f_q = 0.75$ \cite{some_orig_source, leontyev11}, which has been
successfully used to improve the performance of ion force fields \cite{kohagen14,kohagen16,??}.

>>>>

% Here I write pseudo-article statements that will make the main argument.
% Beautiful polished sentences will be formed only afte we agree on these basic things
\section{Introduction}
Cation interactions with cellular membranes play a key role in several biological processes,
like in signal propagation in neurons and vesicle fusion.
Zwitterionic phosphocholine (PC) lipid bilayers have been widely used as model systems
to understand molecular level details of specific cation interactions with
cellular membranes by using experimental \cite{cevc90,tocanne90,binder02,pabst07,uhrikova08}
and theoretical methods \cite{bockmann03,bockmann04,Berkowitz12,??} \todo{We need to select references for these}.
While relative binding affinity of different ions is agreed to follow Hoffmeister
serie, the molecular details of binding and binding energetics are
not fully understood \cite{cevc90,tocanne90,seelig90,binder02,catte16,??}.
%General conclusion from experimental
%studies has been that multivalent ions and lithium have weak specific binding in
%phopholipid bilayers, while other monovalent ions do not essentially
%bind 
%The presence of anoinic lipids, like PS or PG,
%increase the concentration close to the bilayer and thus the amount of bound
%ions, but do not affect the specific binding constant \cite{seelig90}.
Non-invansive spectroscopic methods, like nucelar magnetic resosnance (NMR), scattering
and infrared scpectroscopy, give accurate information about ion binding
in lipid bilayers\cite{hauser76,hauser78,herbette84,akutsu81,altenbach84,binder02}. 
Molecular level interpretation of the results, however, requires assumptions
about binding model and is often not fully conclusive \cite{??}.
On the other hand, the accuracy of the state of the art atomistic resolution lipid and ion models
have turned out insufficient for the detailed interpretation of cation binding details \cite{catte16}.

%The binding details, like binding sites and stoichiometry are not yet fully
%resolved but interpretation of NMR and scattering experiments suggest that one
%Ca2+ interacts mainly with the choline groups \cite{hauser76,hauser78,herbette84} of two
%phospholipid molecules \cite{altenbach84}.

In this work we show that the accuracy of Calcium binding
on zwitterionic PC lipid bilayer can be significantly improved by including electronic polarizability
by using effective continuum correction (ECC)~\cite{leontyev11} for polar region of zwitterionic lipid molecules
in classical MD simulation model.
This is essentially physically well justified version of partial charge scaling
implemented in early days of lipid and surfactant simulations \cite{jonsson86,egberts94}. 
The approach has been previously shown in to improve bulk performace of
ion models against neutron scattering data \cite{Jungwirth2015,kohagen14,kohagen16}.
The better bulk behaviour was not, however, sufficient to correct binding in lipid
bilayers \cite{catte16}. 
%motivation for its application to zwitterionic lipids like POPC.

%Classical molecular dynamics simulations have potential to solve the ion binding details
%in lipid bilayers and reveal its relevance to various biological
%problems \cite{bockmann03,bockmann04,??}. However, the available
%molecular dynamics simulation models have a strong tendency to overestimate
%cation binding on zwitterionic bilayers and none of them can reproduce the
%experimental data with the accuracy required for the interpretation of
%experiments \cite{catte16}.

The proposed MD simulation model with ECC-corrected lipids and ions
reproduce the experimental NMR data of Calcium binding in PC lipid bilayers
and mainly supports the proposed ternary complex model (2 Ca2+ bind in
on POPC) \cite{akutsu81,altenbach84}. However, non-negligible contribution
of one Ca2+ cross bridging three POPC molecules is also observed.
The overestimated specific cation binding observed in current lipid models \cite{catte16}
may lead to articifially positively charged memranes and significant artefacts in
MD simualtions. Thus, the proposed approach to improve the accuracy of cation
binding in current MD models is highly useful for future MD simulations
with physiological salt conditions.

%re is a vast demand for classical molecular dynamics simulation
%model which correctly reproduces cation binding in lipid bilayers.

%The lack of electronic polarizability from the classical MD simulation
%models is a potential source of artefact, which could lead to overbinding
%of cations. 
%Almost all lipid bilayer simulations reported in the literature 
%are done with force fields that represent interactions between atoms
%with pairwise additive empirical potential and exclude the electronic 
%polarizability \cite{SOME_review, FF_papers}.
%The issue has been considered higly relevant since the early days of
%lipid bilayer simulations and pioneering simulation studies scaled
%the partial charges to effectively include electronic
%polarizability \cite{jonsson86,egberts94}.

@@

\subsection{Electronic continuum correction for lipid bilayers}
The explicit inclusion of
electronic polarizability has been, however, turned to be 
practically complicated and implementations for lipids are rare \cite{chowdhary13}.

According to the electronic continuum correction (ECC)\cite{leontyev11}, electronic
polarizability can be included in classical MD simulations by
placing all particles into a homogeneous dielectric continuum 
with a dielectric constant $\epsilon _{el}$, 
which is the electronic part of the dielectric constant of 
the media \cite{leontyev11}. Measurements of high frequency 
dielectric constant gives values of approximately $\epsilon _{el} \approx 2$ 
for almost any biomaterial \cite{some_original_work, leontyev11}. % that exists in living cells. 
Such a continuum can be easily included in standard MD simulation by
a formal transformation of partial charges 
\begin{equation}
  Q^{ECC} = f_q \cdot Q
\end{equation}
with a constant scaling factor $f_q = \epsilon _{el} ^{-1/2}$ 
effectively representing the newly introduced electronic continuum. 
The value measured for water, $\epsilon _{el} = 1.78$, gives 
a scaling factor of $f_q = 0.75$ \cite{some_orig_source, leontyev11}, which has been
successfully used to improve the performance of ion force fields \cite{kohagen14,kohagen16,??}.
